working title Compatibility system and stygma size of pollen recipient as main predictors of heterospecific pollen effect

### 2018-12-04 Manuscript_draft/Manuscript_draft.Rmd
more writing

---
title: Working title, The impact of heterospecific pollen on plant reproductive success is mediated by phylogenetic distance and floral reproductive traits 
author:
- affilnum: '1,2,3'
  name: Jose B. Lanuza, Ignasi Bartomeus, Tia-Lynn Ashman, Romina Rader
  email: barragansljose@gmail.com
affiliation:
- affil: US Environmental Protection Agency, Office of Research and Development, National
    Health and Environmental Effects Research Laboratory, Atlantic Ecology Division,
    27 Tarzwell Drive  Narragansett, RI, 02882, USA
  affilnum: 1
- affil: Big Name University, Department of R, City, BN, 01020, USA
  affilnum: 2
- affil: Estacion Biologica de Donana (EBD-CSIC), E-41092 Sevilla, Spain
  affilnum: 3
output:
  pdf_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    includes:
      in_header: header.tex
    keep_tex: yes
    number_sections: no
    template: manuscript.latex
  word_document: null
  html_document: null
capsize: normalsize
classoption: a4paper
csl: ecology.csl
documentclass: article
fontsize: 11pt
footerdate: yes
geometry: margin=1in
linenumbers: yes
bibliography:
- references.bib
- knitcitations.bib
spacing: doublespacing
abstract: no
---

@@

Recent studies have advanced in the ecological understanding of heterospecific pollen effect @morales2008 @ashman2013 @arceo2016. A general overview of foreign pollen arrival is that it can play an important role on species fitness but seems to be context dependent and not always produce a decrease in fitness @morales2008. Part of this unpredictability is due to the enormous variability of foreing pollen transferred in nature, where levels between 0 and 75 percent are seen, but most commonly values ranges between 10 and 20 percent of the total pollen load, being the generalist species the ones that receive greater loads of heterospecific pollen @montgomery2012 @fang2013. Although heterospecific pollen quantity is fundamental to understand the outcome of the interaction so is the different traits of both pollen donor and recipient. @ashman2013 postulated the first predictive framework for traits of heterospecific pollen effect where traits such as compatibility system and pollen size among others should be crucial to understand foreing pollen effect. Moreover, in @tong2016 a assymetric effect was shown in a crossing experiment between 6 species of the genus *Pedicularis* where the pollen of long styled species was able to grow the ful length of the style on short styled species but not viceversa. Despite these recent caveats, we still lack empirical evidence to affirm what are the main traits that drive heterospecific pollen effect for both pollen donor and recipient at seed production level. Interestingly, this trait based question of effect between a pair of species cannot be solved without consider the phylogenetic relatedness of them. A general view is that close related species will have greater negative impact than far related species (REFS) but few studies test this fact. The interaction between close related species have been tradionally overlooked (Refs) but not always the effect of congeners or confamilias can be seen due to these other species are capable of fertilization (Refs.) Combining both traits and phylogenetic relatedness could give a better picture of how both parts interact in the heterospecific pollen process.

@@

Examples of articles...

can reduce species fitness (REFS) but seems to be highly contex-dependent. There are hypothesized that some traits can play a crucial role in this species interaction such as stigma type, pollen size,

Mention invasive species in this paragraph

Few studies have tried to understand how relatedness is involved in the hp effect but generally 
Until our knowledge ....

Rescue from here the useful things:

Invasive species are supposed to have greater negative effect than native ones @arceo2016. Although when non-natives species don´t have greater negative effect we still don´t know why. For this reason, this ecological question is non a native non native one is a trait based issue that is still to be solved. Moreover, the quantity of pollen that integrates in the network can be quite variable ranging from low quatities @bartomeus2008 to intermediate (ref) to high (ref). Moreover, closely related species are supposed to reduce fitness in greater effect but the evidence is scarce and based on independent studies with different methodologies (Arceo-gomez & Ashman 2016) or studies that just check it with a pair of species that are highly related with the aim to understand hybridization costs (refs). There is a need to deepen into how relatednes is involve in the costs of heterospecific pollen effect. Furthermore, following the conceptual trait framework of Ashman and Arceo-Gomez on heterospecific pollen there are good theoretical basis for trait effect. Notwithstanding, non empirical work has tested how really these traits are involved in heterospecific pollen effect.

Explain traits. Put examples

what is closely related? same genus? Just that right, the rest is far related?

>>>>

---
title: working title Compatibility system and stygma size of pollen recipient as main predictors of heterospecific pollen effect
author:
- affilnum: '1,2,3'
  name: Jose B. Lanuza, Ignasi Bartomeus, Tia-Lynn Ashman, Romina Rader
  email: barragansljose@gmail.com
affiliation:
- affil: US Environmental Protection Agency, Office of Research and Development, National
    Health and Environmental Effects Research Laboratory, Atlantic Ecology Division,
    27 Tarzwell Drive  Narragansett, RI, 02882, USA
  affilnum: 1
- affil: Big Name University, Department of R, City, BN, 01020, USA
  affilnum: 2
- affil: Estacion Biologica de Donana (EBD-CSIC), E-41092 Sevilla, Spain
  affilnum: 3
output:
  pdf_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    includes:
      in_header: header.tex
    keep_tex: yes
    number_sections: no
    template: manuscript.latex
  word_document: null
  html_document: null
capsize: normalsize
classoption: a4paper
csl: ecology.csl
documentclass: article
fontsize: 11pt
footerdate: yes
geometry: margin=1in
linenumbers: yes
bibliography:
- references.bib
- knitcitations.bib
spacing: doublespacing
abstract: no
---

@@

Recent studies have advanced in the ecological understanding of heterospecific pollen effect @morales2008 @ashman2013 @arceo2016. A general overview of foreign pollen arrival is that it can play an important role on species fitness but seems to be context dependent and not always produce a decrease in fitness @morales2008. Part of this unpredictability is due to the enormous variability of foreing pollen transferred in nature, where levels between 0 and 75 percent are seen, but most commonly values ranges between 10 and 20 percent of the total pollen load, being the generalist species the ones that receive greater loads of heterospecific pollen @montgomery2012 @fang2013. Although heterospecific pollen quantity is fundamental to understand the outcome of the interaction so is the different traits of both pollen donor and recipient. @ashman2013 postulated the first predictive framework for traits of heterospecific pollen effect, where different traits such as compatibility system and pollen size among others seems to be crucial to understand foreing pollen effect. Moreover, in @tong2016 a assymetric effect was shown in a crossing experiment between 6 species of the genus *Pedicularis* where the pollen of long styled species was able to grow the full length of the style on short styled species but not viceversa. Despite these recent caveats, we still lack empirical evidence to affirm what are the main traits that drive heterospecific pollen effect for both pollen donor and recipient at seed production level. Interestingly, to comprehend how this traits interact is also crucial to look at the phylogenetic relatedness of the species. There is a considerable amount of literature of crosses between close related species @brown2001 @arceo2016can @tong2016 for which is predicted to have a greater negative impact as @arceo2016 suggest on their work where invasion status and relatedness is considered. Notwithstanding, until our knowledge the effect of related and non-related species remains to be tested explicetely.

@@

The great environmental variability in natural systems and complexity of floral structures make heterospecific pollination studies a daunting task. For this reason, control pollination experiments where hand pollination is applied instead of using pollinators as vectors have simplified this task. However, other issues may arise such as

### 2018-12-04 Manuscript_draft/Manuscript_draft.tex
more writing

\documentclass[11pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[margin=1in]{geometry}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Working title, The impact of heterospecific pollen on plant reproductive success is mediated by phylogenetic distance and floral reproductive traits},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

@@

\title{Working title, The impact of heterospecific pollen on plant reproductive
success is mediated by phylogenetic distance and floral reproductive
traits}
\author{
Jose B. Lanuza, Ignasi Bartomeus, Tia-Lynn Ashman, Romina Rader
}
\date{}
% Allowing for landscape pages
\usepackage{lscape}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}

@@

\begin{document}
%%Edited by @jhollist 09/15/2014
%%Adds title from YAML
\begin{singlespace}
\begin{center}
\huge Working title, The impact of heterospecific pollen on plant reproductive
success is mediated by phylogenetic distance and floral reproductive
traits
\end{center}
%%Adds Author, correspond email asterisk, and affilnum from YAML
\begin{center}
\large
Jose B. Lanuza, Ignasi Bartomeus, Tia-Lynn Ashman, Romina Rader \textsuperscript{*} \textsuperscript{1,2,3} 
\end{center}
%%Adds affiliations from YAML
\begin{justify}
\footnotesize \emph{ 
\\*
\textsuperscript{1}US Environmental Protection Agency, Office of Research and Development,
National Health and Environmental Effects Research Laboratory, Atlantic
Ecology Division, 27 Tarzwell Drive Narragansett, RI, 02882, USA\\*
\\*
\textsuperscript{2}Big Name University, Department of R, City, BN, 01020, USA\\*
\\*
\textsuperscript{3}Estacion Biologica de Donana (EBD-CSIC), E-41092 Sevilla, Spain\\*
}
%%Adds corresponding author email(s) from YAML
\newcounter{num}
\setcounter{num}{1}
\\[0.1cm]
\footnotesize \emph{ 
\ifnum\value{num}=1%
\textsuperscript{*} corresponding author:
\fi
\href{mailto:barragansljose@gmail.com}{\nolinkurl{barragansljose@gmail.com}}
\stepcounter{num}
}
\end{justify}
%%Adds date from YAML
\normalsize

@@

Recent studies have advanced in the ecological understanding of
heterospecific pollen effect Morales and Traveset (2008) Ashman and
Arceo-Gómez (2013) Arceo-Gómez and Ashman (2016). A general overview of
foreign pollen arrival is that it can play an important role on species
fitness but seems to be context dependent and not always produce a
decrease in fitness Morales and Traveset (2008). Part of this
unpredictability is due to the enormous variability of foreing pollen
transferred in nature, where levels between 0 and 75 percent are seen,
but most commonly values ranges between 10 and 20 percent of the total
pollen load, being the generalist species the ones that receive greater
loads of heterospecific pollen Montgomery and Rathcke (2012) Fang and
Huang (2013). Although heterospecific pollen quantity is fundamental to
understand the outcome of the interaction so is the different traits of
both pollen donor and recipient. Ashman and Arceo-Gómez (2013)
postulated the first predictive framework for traits of heterospecific
pollen effect where traits such as compatibility system and pollen size
among others should be crucial to understand foreing pollen effect.
Moreover, in Tong and Huang (2016) a assymetric effect was shown in a
crossing experiment between 6 species of the genus \emph{Pedicularis}
where the pollen of long styled species was able to grow the ful length
of the style on short styled species but not viceversa. Despite these
recent caveats, we still lack empirical evidence to affirm what are the
main traits that drive heterospecific pollen effect for both pollen
donor and recipient at seed production level. Interestingly, this trait
based question of effect between a pair of species cannot be solved
without consider the phylogenetic relatedness of them. A general view is
that close related species will have greater negative impact than far
related species (REFS) but few studies test this fact. The interaction
between close related species have been tradionally overlooked (Refs)
but not always the effect of congeners or confamilias can be seen due to
these other species are capable of fertilization (Refs.) Combining both
traits and phylogenetic relatedness could give a better picture of how
both parts interact in the heterospecific pollen process.

@@

Examples of articles\ldots{}

can reduce species fitness (REFS) but seems to be highly
contex-dependent. There are hypothesized that some traits can play a
crucial role in this species interaction such as stigma type, pollen
size,

Mention invasive species in this paragraph

Few studies have tried to understand how relatedness is involved in the
hp effect but generally Until our knowledge \ldots{}.

Rescue from here the useful things:

Invasive species are supposed to have greater negative effect than
native ones Arceo-Gómez and Ashman (2016). Although when non-natives
species don´t have greater negative effect we still don´t know why. For
this reason, this ecological question is non a native non native one is
a trait based issue that is still to be solved. Moreover, the quantity
of pollen that integrates in the network can be quite variable ranging
from low quatities Bartomeus et al. (2008) to intermediate (ref) to high
(ref). Moreover, closely related species are supposed to reduce fitness
in greater effect but the evidence is scarce and based on independent
studies with different methodologies (Arceo-gomez \& Ashman 2016) or
studies that just check it with a pair of species that are highly
related with the aim to understand hybridization costs (refs). There is
a need to deepen into how relatednes is involve in the costs of
heterospecific pollen effect. Furthermore, following the conceptual
trait framework of Ashman and Arceo-Gomez on heterospecific pollen there
are good theoretical basis for trait effect. Notwithstanding, non
empirical work has tested how really these traits are involved in
heterospecific pollen effect.

Explain traits. Put examples

what is closely related? same genus? Just that right, the rest is far
related?

@@

@@

\hypertarget{ref-bartomeus2008}{}
Bartomeus, I., J. Bosch, and M. Vilà. 2008. High invasive pollen
transfer, yet low deposition on native stigmas in a carpobrotus-invaded
community. Annals of Botany 102:417--424.

@@

>>>>

\documentclass[11pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[margin=1in]{geometry}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={working title Compatibility system and stygma size of pollen recipient as main predictors of heterospecific pollen effect},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

@@

\title{working title Compatibility system and stygma size of pollen recipient
as main predictors of heterospecific pollen effect}
\author{
Jose B. Lanuza, Ignasi Bartomeus, Tia-Lynn Ashman, Romina Rader
}
\date{}
% Allowing for landscape pages
\usepackage{lscape}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}

@@

\begin{document}
%%Edited by @jhollist 09/15/2014
%%Adds title from YAML
\begin{singlespace}
\begin{center}
\huge working title Compatibility system and stygma size of pollen recipient
as main predictors of heterospecific pollen effect
\end{center}
%%Adds Author, correspond email asterisk, and affilnum from YAML
\begin{center}
\large
Jose B. Lanuza, Ignasi Bartomeus, Tia-Lynn Ashman, Romina Rader \textsuperscript{*} \textsuperscript{1,2,3} 
\end{center}
%%Adds affiliations from YAML
\begin{justify}
\footnotesize \emph{ 
\\*
\textsuperscript{1}US Environmental Protection Agency, Office of Research and Development,
National Health and Environmental Effects Research Laboratory, Atlantic
Ecology Division, 27 Tarzwell Drive Narragansett, RI, 02882, USA\\*
\\*
\textsuperscript{2}Big Name University, Department of R, City, BN, 01020, USA\\*
\\*
\textsuperscript{3}Estacion Biologica de Donana (EBD-CSIC), E-41092 Sevilla, Spain\\*
}
%%Adds corresponding author email(s) from YAML
\newcounter{num}
\setcounter{num}{1}
\\[0.1cm]
\footnotesize \emph{ 
\ifnum\value{num}=1%
\textsuperscript{*} corresponding author:
\fi
\href{mailto:barragansljose@gmail.com}{\nolinkurl{barragansljose@gmail.com}}
\stepcounter{num}
}
\end{justify}
%%Adds date from YAML
\normalsize

@@

Recent studies have advanced in the ecological understanding of
heterospecific pollen effect Morales and Traveset (2008) Ashman and
Arceo-Gómez (2013) Arceo-Gómez and Ashman (2016). A general overview of
foreign pollen arrival is that it can play an important role on species
fitness but seems to be context dependent and not always produce a
decrease in fitness Morales and Traveset (2008). Part of this
unpredictability is due to the enormous variability of foreing pollen
transferred in nature, where levels between 0 and 75 percent are seen,
but most commonly values ranges between 10 and 20 percent of the total
pollen load, being the generalist species the ones that receive greater
loads of heterospecific pollen Montgomery and Rathcke (2012) Fang and
Huang (2013). Although heterospecific pollen quantity is fundamental to
understand the outcome of the interaction so is the different traits of
both pollen donor and recipient. Ashman and Arceo-Gómez (2013)
postulated the first predictive framework for traits of heterospecific
pollen effect, where different traits such as compatibility system and
pollen size among others seems to be crucial to understand foreing
pollen effect. Moreover, in Tong and Huang (2016) a assymetric effect
was shown in a crossing experiment between 6 species of the genus
\emph{Pedicularis} where the pollen of long styled species was able to
grow the full length of the style on short styled species but not
viceversa. Despite these recent caveats, we still lack empirical
evidence to affirm what are the main traits that drive heterospecific
pollen effect for both pollen donor and recipient at seed production
level. Interestingly, to comprehend how this traits interact is also
crucial to look at the phylogenetic relatedness of the species. There is
a considerable amount of literature of crosses between close related
species Brown and Mitchell (2001) Arceo-Gómez et al. (2016) Tong and
Huang (2016) for which is predicted to have a greater negative impact as
Arceo-Gómez and Ashman (2016) suggest on their work where invasion
status and relatedness is considered. Notwithstanding, until our
knowledge the effect of related and non-related species remains to be
tested explicetely.

@@

The great environmental variability in natural systems make
heterospecific pollination studies a daunting task. Moreover, when
pollinators are not used as pollen vectors, to control for standarized
pollen quantities hand pollination treatments are applied. However,
different methodologies across studies and different proxys of effect
difficulty comparisons across studies.

@@

\hypertarget{ref-arceo2016can}{}
Arceo-Gómez, G., R. A. Raguso, and M. A. Geber. 2016. Can plants evolve
tolerance mechanisms to heterospecific pollen effects? An experimental
test of the adaptive potential in clarkia species. Oikos 125:718--725.

@@

@@

\hypertarget{ref-brown2001}{}
Brown, B. J., and R. J. Mitchell. 2001. Competition for pollination:
Effects of pollen of an invasive plant on seed set of a native congener.
Oecologia 129:43--49.
